---
title: 'DATA 607: Data Acquisition and Management'
author: "Melvin Matanos, Fall 2022"
subtitle: "Assignment 2"
output:
  word_document: default
  html_document: default
  pdf_document: default
---

## Description

Choose six recent popular movies.  Ask at least five people that you know (friends, family, classmates, imaginary friends) to rate each of these movie that they have seen on a scale of 1 to 5.  Take the results (observations) and store them in a SQL database.  Load the information into an R dataframe.

## Loading the required library

```{r}
library(RMySQL)
```

## Loading required package: DBI

```{r}
library(DBI)
```


```{r connect,message=FALSE}
library(ggplot2)
library(dplyr)
library(grid)
```

# Connect To The Database

```{r}
drv<-dbDriver("MySQL")
con <- dbConnect(drv, user="root", password="Mamamary_0908", host="localhost", port=3306, dbname="moviesdb")
```


## Pulling Data from the Database 

Pull data via the connection.

```{r pull_data}
sql <- 'SELECT `films`.`title`, 
`reviewers`.`name`, 
`ratings`.`rating`
FROM `ratings`
JOIN `films` ON `films`.`id` = `ratings`.`films_id`
JOIN `reviewers` ON `reviewers`.`id` = `ratings`.`reviewers_id`'
movies_df <- dbGetQuery(con, sql)

# Disconnect from the database
disconnected <- dbDisconnect(con)
```

## Exploratory Data Analysis

We have loaded the data into a dataframe. 
Let us perform exploratory data analysis.

```{r first_look}
dim(movies_df)
names(movies_df)
```

```{r}
movies_df[1:10,]
```

```{r}
summary(movies_df$rating)
```

```{r}
movies_df %>%
  group_by(title) %>%
  summarise(min=min(rating),
            mean=mean(rating),
            median=median(rating),
            max=max(rating))
```


Previewbthe data

```{r preview}
head(movies_df)
```

Some of the interesting visualization of my findings:

```{r}
ggplot(movies_df,aes(x=rating))+
  geom_bar(fill="red")+
  ggtitle("Frequency of Ratings")
```

```{r}
ggplot(movies_df, aes(x=title, y=rating))+
  geom_boxplot(fill="LightBlue")+
  ggtitle("Rating of the Movies")
```
  

Except Annihilaton and Deapoint 2 which got the lowest rating the rest of it are in rating 3.



EXAMPLE ONLY:


```{r}
movies_df %>%
  group_by(title) %>%
  summarise(min=min(rating),
            mean=mean(rating),
            median=median(rating),
            max=max(rating))
```


```{r}
movies_df %>%
  group_by(title) %>%
    summarise(min=min(rating),
            mean=mean(rating),
            median=median(rating),
            max=max(rating))
ggplot(movies_df,aes(x=rating))+
         geom_bar(fill="red")+
         labs(x = "Title")
```


```{r}
ggplot(movies_df,aes(x=rating))+
         geom_bar(fill="red")+
         labs(x = "Title")
```

```{r}
ggplot(movies_df, aes(x=title, y=rating, fill=title)) + 
  geom_bar(stat = "identity") + 
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank())
```


```{r}
ggplot(movies_df, aes(x=title, y=rating, fill=title)) + 
  geom_boxplot() + 
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank())
```


