---
title: "DATA 607: Data Acquisition and Management"
author: "Karma Gyatso, Enid Roman, Keeno Glanville, Sangeetha Sasikumar, Melvin Matanos"
subtitle: "Project 3 - What are the most valuable skills?"
date: "2022-10-23"
output:
  html_document: default
  pdf_document: default
---

# Introduction

For Project 3, we were assigned to conduct a thorough analysis of the research question, "What are the most valuable skills" for a data scientist. After researching and evaluating several potential data sources, we have decided to utilize web scraping for analyzing jobs posted on Indeed.com. It is a world-wide job aggregation site and has the highest online traffic in the United States, therefore it is a good representation of the job market. It is one  of the largest job sites in the United States together with Linkedin, Zipcruiter, Craigslist and Monster. Initially, we faced many roadblocks on web scraping but again two heads are better than one. We think that scraping from Indeed gave us a well representation sample to achieve our project goal and answer the research question required. We searched for “data scientist” as the targeted job and searched for jobs within New York. 

We faced many difficulties in performing the scraping process and utilized many links , videos and other resources to extract the data. Time constraint and lack of actual experience defines our weakness in each member of the group. One good thing about our team was all of our communication skills were strong, and we would frequently check in on each other. This was beneficial because we were always on the same page, and we struggled together rather than being lost alone. Although we faced many challenges, we pulled through at the end, with results created. 


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(rvest)
library(readr)
library(tidyverse)
library(dplyr)
library(wordcloud2)
urlclean <- "https://raw.githubusercontent.com/Sangeetha-007/Project-3-607/main/test2.csv"
df1<-read_csv(url(urlclean))

attach(df1)
colnames(df1)
attach(df1)
df<-df1%>%
  select(-c(3:7))%>%
  rename("Job Title" = "keyword")
df
```

We know that within our model we want to explore the various known hard and soft skills need for the roles

```{r}
# Hard Skills
freq_hardskills<- df %>%
    mutate(R = grepl("\\bR\\b,", jobDescription)) %>%
    mutate(python = grepl("Python", jobDescription, ignore.case=TRUE)) %>%
    mutate(SQL = grepl("SQL", jobDescription, ignore.case=TRUE)) %>%
    mutate(hadoop = grepl("hadoop", jobDescription, ignore.case=TRUE)) %>%
    mutate(perl = grepl("perl", jobDescription, ignore.case=TRUE)) %>%
    mutate(matplotlib = grepl("matplotlib", jobDescription, ignore.case=TRUE)) %>%
    mutate(Cplusplus = grepl("C++", jobDescription, fixed=TRUE)) %>%
    mutate(VB = grepl("VB", jobDescription, ignore.case=TRUE)) %>%
    mutate(java = grepl("java\\b", jobDescription, ignore.case=TRUE)) %>%
    mutate(scala = grepl("scala", jobDescription, ignore.case=TRUE)) %>%
    mutate(tensorflow = grepl("tensorflow", jobDescription, ignore.case=TRUE)) %>%
    mutate(mongodb = grepl("mongodb", jobDescription, ignore.case=TRUE)) %>%
    mutate(Hive = grepl("Hive", jobDescription, ignore.case=TRUE)) %>%
    mutate(tableau = grepl("tableau", jobDescription, ignore.case=TRUE)) %>%
    mutate("Power BI" = grepl("Power BI", jobDescription, ignore.case=TRUE)) %>%
    mutate(noSQL = grepl("noSQL", jobDescription, ignore.case=TRUE)) %>%
    mutate("predictive modeling" = grepl("predictive modeling", jobDescription, ignore.case=TRUE)) %>%
    mutate(AWS = grepl("AWS", jobDescription, ignore.case=TRUE)) %>%
    mutate(Azure = grepl("Azure", jobDescription, ignore.case=TRUE)) %>%
    mutate(javascript = grepl("javascript", jobDescription, ignore.case=TRUE)) %>%
    mutate(spark = grepl("spark", jobDescription, ignore.case=TRUE)) %>%
    mutate("Machine Learning" = grepl("Machine Learning", jobDescription, ignore.case=TRUE)) %>%
    mutate(cloud = grepl("cloud", jobDescription, ignore.case=TRUE)) %>%
    mutate(masters = grepl("masters", jobDescription, ignore.case=TRUE)) %>%
    mutate(statistics = grepl("statistics", jobDescription, ignore.case=TRUE)) %>%
  
select(`Job Title`, R, python, SQL, hadoop, perl, matplotlib, Cplusplus, VB, java, "Machine Learning", scala, tensorflow, mongodb, javascript, spark, Hive, tableau, "Power BI", noSQL,"predictive modeling", AWS, Azure, cloud, masters, statistics)
freq_hardskills


#Soft Skills
freq_softskills <- df%>%
    mutate(remote = grepl("remote", jobDescription, ignore.case=TRUE)) %>%
    mutate(communication = grepl("communicat", jobDescription, ignore.case=TRUE)) %>%
    mutate(collaborative = grepl("collaborat", jobDescription, ignore.case=TRUE)) %>%
    mutate(creative = grepl("creativ", jobDescription, ignore.case=TRUE)) %>%
    mutate(critical = grepl("critical", jobDescription, ignore.case=TRUE)) %>%
    mutate(problemsolving = grepl("problem solving", jobDescription, ignore.case=TRUE)) %>%
    mutate(activelearning = grepl("active learning", jobDescription, ignore.case=TRUE)) %>%
    mutate(hypothesis = grepl("hypothesis", jobDescription, ignore.case=TRUE)) %>%
    mutate(organized = grepl("organize", jobDescription, ignore.case=TRUE)) %>%
    mutate(judgement = grepl("judgement", jobDescription, ignore.case=TRUE)) %>%
    mutate(selfstarter = grepl("self Starter", jobDescription, ignore.case=TRUE)) %>%
    mutate(interpersonalskills = grepl("interpersonal skills", jobDescription, ignore.case=TRUE)) %>%
    mutate(attentiontodetail = grepl("attention to detail", jobDescription, ignore.case=TRUE)) %>%
    mutate(visualization = grepl("visualization", jobDescription, ignore.case=TRUE)) %>%
    mutate(motivated = grepl("motivated", jobDescription, ignore.case=TRUE)) %>%
    mutate(independent = grepl("independent", jobDescription, ignore.case=TRUE)) %>%
    mutate(resourceful = grepl("resourceful", jobDescription, ignore.case=TRUE)) %>%
    mutate(passion = grepl("passion", jobDescription, ignore.case=TRUE)) %>%
    mutate(determination = grepl("determination", jobDescription, ignore.case=TRUE)) %>%
    mutate(focus = grepl("focus", jobDescription, ignore.case=TRUE)) %>%
    mutate(leadership = grepl("leadership", jobDescription, ignore.case=TRUE)) %>%

  select(`Job Title`, remote, communication, collaborative, creative, critical, problemsolving, 
  activelearning, hypothesis, organized, judgement, selfstarter, interpersonalskills, attentiontodetail, 
  visualization, leadership, motivated, resourceful, independent, passion, determination, focus)
freq_softskills
```


# Analysis
### A summary of the various Hard and Soft skills needs to be established in order for the following visualization tool to utilize the data
```{r}
hardskills <- freq_hardskills %>%
  select(-1)%>%
  summarise_all(sum)%>%
  gather(skill, freq)%>%
  arrange(desc(freq))

softskills <- freq_softskills %>%
  select(-1)%>%
  summarise_all(sum)%>%
  gather(skill, freq)%>%
  arrange(desc(freq))

hardskills
softskills

```
  Top 3 hard skills : Python, SQL, Machine Learning    
  Top 3 soft skills: Communicative, Collaborative, Focused



# Visualization of Soft and Hard Skills

```{r}

ggplot(hardskills,aes(x=reorder(skill, freq), y=freq)) + geom_bar(stat='identity',fill="blue") + xlab('') + ylab('Frequency') + labs(title='Hard Skills') + coord_flip() + theme_minimal()



ggplot(softskills,aes(x=reorder(skill, freq), y=freq)) + geom_bar(stat='identity',fill="purple") + xlab('') + ylab('Frequency') + labs(title='Soft Skills') + coord_flip() + theme_minimal()

```



```{r}
ggplot(hardskills, aes(x=skill, y=freq))+
  geom_point()+
  theme(axis.text.x = element_text(angle = 90))

ggplot(softskills, aes(x=skill, y=freq))+
  geom_point()+
  theme(axis.text.x = element_text(angle = 90))

```


# Conclusion


Our project demonstrates that skills in Python, SQL and Machine Learning are among the top skills required to be considered as an ideal candidate for a Data Scientist roles.This is significant because while these are the core skills, their are various other hard skills that could be utilized to set someone apart from the fellow candidates.It is also Important to note that there are also the soft skills of being communicative, collaborative and focused that also bring an added benefit towards a candidate. Another important thing to note is that the magnitude of hard skills outweighs
the soft skills and thus can demonstrate that more importance is placed on candidates abilities to accurately draw insights from data.


# Enjoy!

```{r}
wordcloud2(hardskills, size = 0.7)
```


















































