---
title: 'DATA 607: Data Acquisition and Management'
author: "Melvin Matanos, Fall 2022"
subtitle: "Week 6 - Assignment"
output:
  html_document: default
  pdf_document: default
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Description

The task for this week is to load an "untidy" .csv file and transofrm it into a tidy dataset. It is containing a data for the two airines. After data wrangling, an anlysis will be done accordingly by comparing the arrivals delays of the two airlines. 


# Loading all the required libraries

```{r}
library(tidyr)
library(dplyr)
library(plyr)
library(ggplot2)
```

# Reading the data from Github 

A .csv file was created that is exactly the same to the assignment data and upload it to my githu repository. Read the data from .csv into a dataframe.


```{r}
#read in data from GitHub
untidy_Flight_csv <- read.table("https://raw.githubusercontent.com/melvinmatanos2008/data_607_fall_2022/main/flights.csv",header=TRUE,sep=",")

```


```{r}
#Visualize csv data
untidy_Flight_csv

```


We need to rename the columns then remove the empty row(s). 
Let us fill in the airline name for the records of delayed flights.

```{r}
tidy_Flight_csv <- dplyr::rename(untidy_Flight_csv, "AirLine" = colnames(untidy_Flight_csv)[1], "status" = colnames(untidy_Flight_csv)[2])
tidy_Flight_csv <- tidy_Flight_csv %>%
  filter(status != "") 
tidy_Flight_csv$AirLine[seq(2,length(tidy_Flight_csv$status),2)] <- tidy_Flight_csv$AirLine[seq(1,length(tidy_Flight_csv$status),2)]
tidy_Flight_csv

```

By using the gather function in order to transform the table to be tidy.

```{r}
tidy_Flight_csv <- tidy_Flight_csv %>% 
  gather(destination, num_of_flights, 3:7)
tidy_Flight_csv
```

Below are the following verall delay rate or the two airlines.

```{r}
ddply(tidy_Flight_csv, "AirLine", transform, perct = num_of_flights / sum(num_of_flights)) %>% 
  filter(status == "delayed") %>%
  group_by(AirLine) %>% 
  dplyr::summarise(total_perct = sum(perct))
```


There is not much significant large difference between the two airlines.

Displayed the overall delay rate of the five destinations.

```{r}
ddply(tidy_Flight_csv, "destination", transform, perct = num_of_flights / sum(num_of_flights)) %>% 
  filter(status == "delayed") %>%
  group_by(destination) %>% 
  dplyr::summarise(total_perct = sum(perct))
```


There is a quite remarkable observation of range of the rates.


# Analysis

The graph shows the total number of flights by airline and the destination alon with the total number of delayed flights by airline and destination.

```{r}
tidy_Flight2_csv <- ddply(tidy_Flight_csv, c("AirLine", "destination"), transform, total_flights =  sum(num_of_flights)) %>% 
  filter(status == "delayed")

ggplot(data = tidy_Flight2_csv, aes(x = paste(destination,"\n",AirLine))) +
  theme(axis.text.x = element_text(angle = 90)) +
  xlab("Air Line and Destination") +
  ylab("Total Number of Flights") +
  geom_bar(aes(y=total_flights),stat = "identity",col = "red", fill = "red") +
  geom_bar(aes(y=num_of_flights),stat = "identity",col = "blue", fill = "blue") +
  geom_rect(mapping=aes(xmin=9.5, xmax=10, ymin=5000, ymax=5250),col = "red", fill = "red") +
  geom_rect(mapping=aes(xmin=9.5, xmax=10, ymin=4500, ymax=4750),col = "blue", fill = "blue") +
  geom_text(mapping=aes(x=9.75, y=5250, label="Total Num of Flight"), size=3, vjust=3, hjust=0.5) +
  geom_text(mapping=aes(x=9.75, y=4750, label="Total Num of Delayed"), size=3, vjust=3, hjust=0.5)
```

In order for us to have a quick easier comparsion,let us compute the percentage of delays by airline and destinations.

```{r}
tidy_Flight2_csv <- mutate(tidy_Flight2_csv, delay_perct = num_of_flights / total_flights)
tidy_Flight2_csv
```
Lets make some descriptive statistic to compare the two airlines

```{r}
print("ALASKA")
```
```{r}

tidy_Flight2_csv %>% 
  filter(AirLine == "ALASKA") %>% 
  select(delay_perct) %>% 
  summary()
```


```{r}
print("AM WEST")
```


```{r}
tidy_Flight2_csv %>% 
  filter(AirLine == "AM WEST") %>% 
  select(delay_perct) %>% 
  summary()
```


Based on the above statistics it is clearly showing that the AM WEST is higher than the ALASKA arlines. Therefor we can conclude that the ALASKA airlines has a better performance in making their flights on time.







