---
title: 'DATA 607: Data Acquisition and Management'
author: "Melvin Matanos, Fall 2022"
subtitle: "Assignment 3"
output:
  html_document: default
  word_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# Loading the reqiured libraries.

```{r}
library(tidyverse)
library(stringr)
library(readr)
```


# Question 1

Using the 173 majors listed in fivethirtyeight.com’s College Majors dataset [https://fivethirtyeight.com/features/the-economic-guide-to-picking-a-college-major/], provide code that identifies the majors that contain either “DATA” or “STATISTICS”

# Answer:

## Loading the data 

```{r}
df_majors <- read.csv(url('https://raw.githubusercontent.com/fivethirtyeight/data/master/college-majors/majors-list.csv'), stringsAsFactors = F)
str(df_majors)
```


# Majors containing data or statistics

```{r}
df_majors$Major[grepl("DATA", df_majors$Major)]
```

```{r}
df_majors$Major[grepl("STATISTICS", df_majors$Major)]
```

# Question 2

Write code that transforms the data below:

[1] “bell pepper” “bilberry” “blackberry” “blood orange”

[5] “blueberry” “cantaloupe” “chili pepper” “cloudberry”

[9] “elderberry” “lime” “lychee” “mulberry”

[13] “olive” “salal berry”

Into a format like this:

c(“bell pepper”, “bilberry”, “blackberry”, “blood orange”, “blueberry”, “cantaloupe”, “chili pepper”, “cloudberry”, “elderberry”, “lime”, “lychee”, “mulberry”, “olive”, “salal berry”)

# Answer:

```{r}
df_fruits_list <- '[1] "bell pepper"  "bilberry"     "blackberry"   "blood orange" [5] "blueberry"    "cantaloupe"   "chili pepper" "cloudberry"  [9] "elderberry"   "lime"         "lychee"       "mulberry"    [13] "olive"        "salal berry"'
```

```{r}
df_fruits_pattern <- str_extract_all(df_fruits_list,pattern = '[A-Za-z]+.?[A-Za-z]+')
df_fruits_list_vector <- str_c(df_fruits_pattern, collapse = ", ")
```


```{r}
writeLines(df_fruits_list_vector)
```

# Question 3

Describe, in words, what these expressions will match: 

(.)\1\1

# Answer:

This will match to the same character appearing three times in a row.

For example: 

```{r}
w_txt <- c("melllllvin","melvvvvvvin","melvin")
str_view(w_txt,"(.)\\1\\1")
```


“(.)(.)\2\1”

# Answer:

This will match something like a 4 letters palindrome.A pair of characters followed by the same pair of characters in reversed order.
For example: 
 
```{r}
w_txt2 <- c("woowing","daamn","zoom","lool")
str_view(w_txt2,"(.)(.)\\2\\1")
```


(..)\1

# Answer:

This will match to any two characters repeated.

For example:

```{r}
w_txt3 <- c("mahahand","sasasand","paparazzi","contigo","dada")
str_view(w_txt3,"(..)\\1")
```


“(.).\1.\1”

# Answer:

This will match to a character followed by any character, the original character, any other character, the original character again. For example: "alaba", "c8c.c".


```{r}
w_txt4 <- c("amana","eleven","event","avatar")
str_view(w_txt4,"(.).\\1.\\1")
```



"(.)(.)(.).*\3\2\1"

# Answer:

This will match to three characters followed by zero or more characters of any kind followed by the same three characters but in reverse order. For example "abccba" or "abc1cba".

```{r}
w_txt5 <- c("sadisdas","madisnotbdam","layisyal","soonisnoos","docanbeod")
str_view(w_txt5,"(.)(.)(.).*\\3\\2\\1")
```

# Question 4

# Construct regular expressions to match words that:

a. Start and end with the same character.

# Answer:
 
 "(.).*\1"

```{r}
str_view(c("remember","level","wind","trust"),"(.).*\\1")
```


b. Contain a repeated pair of letters (e.g. “church” contains “ch” repeated twice.)

# Answer:

"(..).*\1"

```{r}
str_view(c("therefore","require","decide","right"),"(..).*\\1")
```
c. Contain one letter repeated in at least three places (e.g. “eleven” contains three “e”s.)

# Answer:

 “(.).\1.\1”
 
```{r}
str_view(c("available","expense","business","math"),"(.).*\\1.*\\1")
```
 
